{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2020-11-29 documents/acknowledgement.tex
Fluor. working great
Wording
No analytic solution

Prof. Griffiths and Dr. Brovko at UGuelph, Dr. Anany at the GRDC for tips on the Beta-Glo assay

@@

Profs David and Taylor (among others) at YorkU for being super nice, and H.P. and ice9 for tolerating ramblings.

Thanks to L.H. for wisdom regarding this project;

@@

R.I.P. to Dr. Robert Adair, whose work features prominently here, who died Sep. 2020.

@@

The authors reluctantly acknowledge the Computational Electromagnetics Group at The University of Texas at Austin for going through the trouble to develop the AustinMan and AustinWoman human body models and then making them unavailable at https://sites.utexas.edu/austinmanaustinwomanmodels/, and also not replying to email.

It's okay, I'm not bitter.

@@

The authors gratefully acknowledge the IT.IS foundation for the same, and also compiling the tissue property databases.

>>>>

Prof. Griffiths and Dr. Brovko at UGuelph, Dr. Anany at the GRDC for useful tips on the Beta-Glo assay.

@@

Profs H. Kassiri, C. David and W. Taylor (among others) at YorkU for being super nice, and H.P. and ice9 for tolerating ramblings.

Thanks to L.H. for wisdom regarding this project.

@@

R.I.P. to Dr. Robert Adair, whose work features prominently here, died Sep. 2020.

@@

The authors gratefully acknowledge the IT.IS foundation

and also compiling the tissue property databases.

@@

Thanks to Biotium Inc for support on GelGreen.

### 2020-11-29 documents/application.tex
No analytic solution

SARS is found widely distributed throughout many of the most favoured organs \cite{Organ2004}, shielded by an average of 4 cm of chest wall [Schroeder 2013].

\cite{neuroinvasive2020} \cite{situ2020}

However, it would be an understatement to say that destruction of lung tissue is significant in the lethality of SARS [Nicholls 2006].

@@

Applying microwave power to the body is already used extensively in diathermy (the internal heating of tissue for therapeutic effect) and hyperthermy (the overheating or cauterization of tissue for destructive effect). While the majority of diathermy systems use much lower frequencies for deeper heating penetration, several papers have explored the 10 GHz band. Superficial vs deep-tissue hyperthermia.

@@

In addition to these limits, the "microwave hearing threshold" - the pulse energy density threshold at which nerve stimulation begins to occur, should be respected.

>>>>

Free-space "vaccination" of people or spaces is all well and good; but a treatment would seem to be superior. Far fewer installations would be required, cost would be of less object, and effectiveness would not depend on widespread public acceptance. Safety concerns due to continuous exposures would also be lessened.

And so, is it possible to non-invasively produce this effect deep within infected tissue - to break every capsid in the body wholesale?

There is much historical precedent for applying microwave power to the body diathermy (the internal heating of tissue for therapeutic effect) and hyperthermy (the overheating or cauterization of tissue for destructive effect).

While the majority of diathermy systems use much lower frequencies for deeper heating penetration, several papers have explored the 10 GHz band, almost exclusively for superficial hyperthermia.

NCoV has a unique appreciation for nervous system \cite{neuroinvasive2020}, and SARS is usually found widely distributed throughout many of the most well-regarded organs \cite{Organ2004}, shielded by an average of 4 cm of chest wall [Schroeder 2013].

However, it would be an understatement to say that destruction of lung tissue is significant in the lethality of SARS [Nicholls 2006], \cite{situ2020}; largely the lung type II alveolar cells

@@

@@

In addition to these limits, the "microwave hearing threshold" - the pulse energy density threshold at which the slightest nerve stimulation begins to occur, should also be respected. In fact, ultra-high pnerve stimulation has been

### 2020-11-29 documents/background.tex
NLTL doesn't seem to help.
No analytic solution

we appear to have lost confidence in the concept of accurate measurement itself.

@@

@@

\subsection{Artifacts in past absorption spectroscopy experiments, and the subsequent development of non-classical damping hypotheses}

@@

crude, amateurish equipment, by making completely justifiable changes in calibration and post-processing correction for thermal drift, normalization, etc, we could produce a resonance mode at many frequencies. almost any frequency we liked. We were able to produce many candidate peaks even in pure water.

autem We have not yet examined the details of the hydration-layer theory.

@@

However, as Foster et al demonstrated, there is nothing necessarily wrong with VNA-based dielectric spectrometry, if it is well conducted. The Liu and Yang group seem very competent experimenters, and the experiments are very well conducted; but history would seem to suggest that we take some degree of caution.

>>>>

@@

Most things with a stiffness matrix has some resonant frequency\citationneeded. As far as we can tell, two justifications underpin the categorical implausibility of resonance-like behavior in

\begin{itemize}
		\item Viscous damping from biological solvents is so strong that the relaxation time is 
\end{itemize}

To a first order, this simply doesn't seem to be true in the case of the virion.

\begin{itemize}
		\item The charge 
\end{itemize}

@@

\subsection{Artifacts in past absorption spectroscopy experiments, and the subsequent development and debunking of non-classical damping hypotheses}

@@

In the case of our crude, amateurish equipment, by making completely justifiable changes in calibration and post-processing correction for thermal drift, normalization, etc, we could produce a resonance mode at almost any frequency we liked. We were able to produce many candidate peaks even in pure water.

@@

However, as Foster et al demonstrated, there is nothing necessarily wrong with VNA-based dielectric spectroscopy if it is well conducted. The Liu and Yang group seem very competent experimenters, and the experiments are very well conducted; but history would seem to suggest that we take some degree of caution.

### 2020-11-29 documents/charge.tex
osmotically-stressed phage doesn't do it either
No analytic solution

%!TeX root = charge
\documentclass[paper.tex]{subfiles}
\begin{document}

@@

@@

Is such a charge plausible?

As might be expected, the electrostatic properties of the virus differ from that of host cells, both on the sub-protein level\cite{Icosahedral2019} \footnotemark \ and in the lipid bilayer \cite{Lipid2015} \footnotemark. The mechanism given is that ions from solution counteract the protein charges to form an "electric double layer". In any case, the best representation is two concentric spheres with a net charge each.

\footnotetext{
	\cite{Icosahedral2019}: \begin{quote}
		Different from insect, plant, and human, viruses had more positively charged than negatively charged segments. Viral proteins were the only class enriched in extremely high positively charged segments (charge greaterthan + 17) when compared to the overall proteome (1B-inset with p-values).
	\end{quote}
}

@@

@@

\subsection{Major proteins and other components}

Protein net charges extracted from primary protein sequences compiled by the proteome-pI database \cite{ProteomepI2017} and UniProt using localCIDER \cite{CIDER2017}. See \ghfile{biology/data/net_charge.py} and structure.csv. Detailed source information will be added later! There are many.

Note that this is considering the net charge of a whole isolated protein. Some proteins poke through the envelope; if a zwitterionic polarization existed from one end to another, for instance, this would not count here but would affect the coupling to an EM field. Given the size of most proteins, however, this doesn't seem like a likely effect.

@@

$$ U = \frac{3}{5}  \frac{1}{4 \pi \epsilon_0} \frac{Q^2}{R} \approx 2.3\text{e}^{-7} \text{ J} $$
%U = (3/5) * 1/(4*pi*electricConstant) *((1e7 electricCharge)^2 / 60 nanometers) = 2.30708e-7 J

The total is $6 \times 10^{11}$ ATP-equivalent.

@@

Influenza is approximately 174 MDa Ruigrok. Assuming a typical 50\% H concentration, such a charge would require about 1 in every 15 atoms to have a free electron. Compare to less than 1/100 for Arginine or 1/600 for DNA.

@@

\footnotetext{
For an impermeable virus like the even-numbered T's, the inside salt concentration is set by the incubation solution's value of nb. Anderson et al. (1953) found in the case of T6, upon rapid dilution with distilled water, for example, that incubation nb values on the order of 1.5 M were required for osmotic shock. From Eq. 2, and n0 \approx 3 M, it follows that nin \approx 0.62 M and hence n++ n- \approx 4.2 M, which because of the Donnan equilibrium is indeed significantly larger than 2nb \approx 3 M.
}
https://www-sciencedirect-com.ezproxy.library.yorku.ca/science/article/pii/S0006349503744555

@@

If we assume that the charge is not as. All other properties being equal, to account for the same coupling and absorption, one of the following would seem to need to be true:

1. the effective path length of the waveguide used by Sun may not be as they expected.
2. The concentration of virus was higher than measured.

On the largest scale, at a pH of 7, the net charges of the 11 proteins coded for by the Influenza A genome only contribute

>>>>

%!TeX root = charge
\documentclass[paper.tex]{subfiles}
\usepackage{parselines}

\begin{document}

@@

\begin{autem}
	What is the charge on phage x174 from Burkhauzens Raman spectroscopy data?
\end{autem}

@@

The mechanism given is that ions from solution counteract the protein charges to form an "electric double layer". In any case, the best representation is two concentric spheres, with a net charge each.

What specifically contributes to this charge? For the purposes of coarse MD simulation, it seemed useful to localize the charge concentration.

As might be expected, the electrostatic properties of the virus differ from that of host cell, both on the sub-protein level\cite{Icosahedral2019} \footnotemark \ and in the lipid bilayer \cite{Lipid2015} \footnotemark.

\footnotetext{
	\cite{Icosahedral2019}: \begin{quote}
		Different from insect, plant, and human, viruses had more positively charged than negatively charged segments. Viral proteins were the only class enriched in extremely high positively charged segments (charge $> +17$) when compared to the overall proteome (1B-inset with p-values).
	\end{quote}
}

@@

\footnote{Influenza also has several other fascinating electrostatic properties. First, the net charge on the virion changes}

@@

\subsection{Major components}

At the largest scale, at a pH of 7 the net charges of the 4 proteins that comprise ~88\% of the protein mass\cite{Quantitative1981} of the virus (of the 11 coded for by the Influenza A genome) only contribute at most  $\approx 1 \times 10^5\  e^-$.

\begin{sidenote}
Protein net charges extracted from protein primary structure sequences\footnotemark compiled by the proteome-pI database \cite{ProteomepI2017} and UniProt\footnotemark using localCIDER \cite{CIDER2017}. See \ghfile{biology/data/net_charge.py} and structure.xlsx. Detailed source information will be added later!
\end{sidenote}
Assuming a charge of $2e^-$ per base pair due to phosphate groups, the 13.5 kbp of RNA can hardly contribute $> 3e^- \times10^4$

Note that this is considering the net charge of a whole isolated protein. Some proteins poke through the envelope; if a zwitterionic polarization existed from one end of a protein to another, for instance, this would not count here but would affect the coupling to an EM field. Given the small extent of most of the structural proteins, however, this doesn't seem like a likely effect.

\begin{sidenote}
	It should be noted that these are the simplest, first-order primary structure charges. In practice, the charge is determined by the secondary structure and the bonding to adjacent structures.

	 Penhoet \cite{Structurea}, "the viral glycoproteins and nonstructural protein are heterogeneous in isoelectric point, forming constellations of interrelated protein spots." whether this is intra- or inter-virion hetrogenity is unknown to us.

	Empirically-determined charges appear to be somewhat inconsistent:

	Privalsky and Penhoet \cite{Influenza1978} and Penhoet \cite{Structurea} find that the isoelectric point of the M protein is so acidic that it runs right off the end of their 2D electrophoresis gel. They also note that

	However, the predicted value by [] is about 3.

	This is not of much importance for this section, but may be important when setting up MD simulations.

	Because CryoEM uses charged particles, it is theoretically possible to extract the actual charge distribution of a macrostructure directly from an un-fit CryoEM map \cite{Identification2018}. 
\end{sidenote}

@@

$$ U = \frac{3}{5}  \frac{1}{4 \pi \epsilon_0} \frac{Q^2}{R} \approx 2.3\text{e}^{-7} \text{ J} \approx 6 \times 10^{11}\  \text{ATP-equivalent} $$
%U = (3/5) * 1/(4*pi*electricConstant) *((1e7 electricCharge)^2 / 60 nanometers) = 2.30708e-7 J

@@

Influenza is approximately 174 MDa Ruigrok. Assuming a typical 50\% H concentration, such a charge would require about 1 in every 15 atoms to have a free electron - a very different definition of "blood plasma". Compare to less than 1/100 for Arginine or 1/600 for DNA.

@@

\footnotetext{
For an impermeable virus like the even-numbered T's, the inside salt concentration is set by the incubation solution's value of nb. Anderson et al. (1953) found in the case of T6, upon rapid dilution with distilled water, for example, that incubation nb values on the order of 1.5 M were required for osmotic shock. From Eq. 2, and n0 \approx 3 M, it follows that nin $\approx 0.62 M$ and hence n++ n- \approx 4.2 M, which because of the Donnan equilibrium is indeed significantly larger than 2nb \approx 3 M.
}
https://www-sciencedirect-com.ezproxy.library.yorku.ca/science/article/pii/S0006349503744555

@@

\subsection{Discussion}

If we assume that the charge is not as. All other properties being equal, to account for the same coupling and absorption, it seems one or many of the following would seem to need to be true:

\begin{itemize}
\item We have grossly misunderstood the nature of the charge imbalance. Most likely.
\item The effective path length of the waveguide used by Sun may not be as they expected, or otherwise the corrections made are not of the correct scale.
\item The concentration of virus was higher than measured by plaque assay - perhaps a dilution was missed.
\end{itemize}
